---
title: "The implementation of Bayesian Networks in African recessions analysis"
author: "Lorenzo Rossi"
date: "24/04/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In 2003-2007, the developing world experienced an impressive economic boom, growing
at a rate of 7% per year. The boom was fueled by a mix of four ingredients prevailing in
global markets: exceptional financing, high commodity prices and, for a significant number
of countries, large flows of remittances. The first two conditions had coincided for the last
time in the 1970s, while the mix of the three had never been experienced before. The rise of
an alternative Asian engine, with China at the center, is a fourth element, which has had a
strong influence on world trade and commodity prices.


#######
ON BN

Bayesian networks are graphical models where nodes represent random variables (the two
terms are used interchangeably in this article) and arrows represent probabilistic dependencies
between them (Korb and Nicholson 2004).

The main role of the network structure is to express the conditional
independence relationships among the variables in the model
through graphical separation

The graphical structure $G = (V, A)$ of a Bayesian network is a directed acyclic graph (DAG),
where V is the node (or vertex ) set and A is the arc (or edge) set. The DAG defines a
factorization of the joint probability distribution of $V = {X1, X2, . . . , Xv}$, often called the
global probability distribution, into a set of local probability distributions, one for each variable.
The form of the factorization is given by the Markov property of Bayesian networks (Korb and
Nicholson 2004), which states that every random variable $Xi$ directly depends
only on its parents $ΠXi$
:


#ON PC ALG
Understanding cause-effect relationships between variables is of primary interest in many fields
of science. Usually, experimental intervention is used to find these relationships. In many
settings, however, experiments are infeasible because of time, cost or ethical constraints.
More precisely, the causal effect C(V1, V6, x˜) of V1 from V1 = ˜x on V6 is defined as... 
If the causal relationships are linear, the causal effect is independent of ˜x.
Since the causal structure was not identified uniquely in our example, we cannot expect to get a unique number for the causal effect

Let G = (V, E) be a graph consisting of a set of vertices V
and a set of edges E ⊆ V × V. The set of vertices that are
adjacent to A in graph G is defined as: adj(A, G) = {B :
(A, B) ∈ E or(B, A) ∈ E}. B is called a collider if we have
the v-structure: A → B ← C.
Graph G is a Directed Acyclic Graph (DAG) if G contains
only directed edges and has no directed cycles. The skeleton
of a DAG G is the undirected graph obtained from G by
ignoring the direction of the edges. An equivalence class of
DAGs is the set of DAGs which have the same skeleton
and the same v-structures. An equivalence class of DAGs
can be uniquely described by a completed partially directed
acyclic graph (CPDAG) which includes both directed and
undirected edges.

The first step of the algorithm consist of to build the skeleton graph: for each pair of vertices u and v,

It assigns the directions to the edges in the skeleton, using the list of conditional independencies found during the Step 1. The output is completed partially directed acyclic graph, a mixed graph where there are both directed and bidirected arcs.

The PC algorithm [2] (original-PC algorithm henceforward)
has two main steps. In the first step, it learns from data a
skeleton graph, which contains only undirected edges. In
the second step, it orients the undirected edges to form
an equivalence class of DAGs. As the first step of the PC
algorithm contributes to most of the computational costs, we
only focus on the modification of this skeleton learning step
in this paper, and information about the edge orientation
step can be found in [2].
The theoretical foundation of the original-PC algorithm
[2] is that if there is no link (edge) between nodes X and Y ,
then there is a set of vertices Z that either are neighbours of
X or Y such that X and Y are independent conditioning on
Z. In other words, Z disconnects X and Y

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

#EConomic Background 

A recession (fall in national income) will typically be characterised by high unemployment, falling average incomes, increased inequality and higher government borrowing. The impact of a recession depends on how long it lasts and the depth of the fall in output.

The main costs of a recession will be:

Unemployment
Fall in income – shorter working week.
Rise in poverty
Fall in asset prices (e.g. fall in house prices/stock market)
Increased inequality and an increase in relative poverty
Higher government borrowing (less tax revenue)
Permanently lost output.
Firms go out of business.

In particular, aggregate productivity can be affected by recessions in two separate manners: i) by means of their impact on productivity within each sector; ii) by inducing sectoral reallocations of input factors across different sectors.4 The effect through sectoral reallocation is not clear, since labor can move between various low- and high-productivity sectors, with an ambiguous net effect on productivity. Counter-cyclical reallocation takes place when input factor reallocation during a downturn leads to less productive jobs being destroyed and labor moving into more productive uses (Mortensen and Pissarides, 1994). Pro-cyclical reallocation occurs when more productive industries are disproportionately affected by recessions, for example due to credit constraints (Barlevy, 2003).

Therefore, the extent to which recessions impact total factor productivity (TFP) is ultimately an empirical question

##ON AFRICA

Years Covered: 2000 to 2017.

Countries Covered: 27 African Countries Including: Morocco, South Africa, Tanzania, Rwanda, Eswatini, Togo, Burkina Faso, Angola, Tunisia, Nigeria, Kenya, Burundi, Benin, Namibia, Central African Republic, Sudan, Gabon, Niger, Sierra Leone, Lesotho, Mauritania, Senegal, Mauritius, Botswana, Cameroon, Zimbabwe and Mozambique.

Dramatic increases in international agricultural commodity prices began in 2006 and peaked in July 2008. An equally remarkable and rapid decline of those prices then ensued, accompanied by extreme volatility in those prices. The trend in food prices lagged the rapid increases in other commodity prices, including oil and metals, but accompanied those other prices in the downward part of the cycle.

Greater regional integration offers more opportunities to climb the technological
ladder than do exports outside the continent, given that intra-African exports are
technologically more advanced, as indicated by a larger share of medium- and
high-technology manufactures (International Trade Centre and UNCTAD, 2021; Saygili
et al., 2018). Moreover, intra-African trade consists of a higher share of processed
goods (41 per cent) than exports to the rest of the world (17 per cent), and it is
much more diversified in terms of traded products (International Trade Centre and
UNCTAD, 2021).

Obviously the collapse of the stock exchanges in the great finance centres in May 2008 was also promptly transmitted to the stock exchanges in the most important emerging countries. 
Consequently, not only were the flows of portfolio and direct investments to the developing countries significantly lower, but commercial bank credits and non-bank financing were also reduced. On account of the world economic recession, there was a sharp fall in the demand for goods and services from the developing and emerging countries.

Dramatic increases in international agricultural commodity prices began in 2006 and peaked in July 2008. An equally remarkable and rapid decline of those prices then ensued, accompanied by extreme volatility in those prices. The trend in food prices lagged the rapid increases in other commodity prices, including oil and metals, but accompanied those other prices in the downward part of the cycle. n developing countries, poor market integration and border barriers may have limited pass-through of these prices to the farmgate, but there was more rapid food price and general inflation than occurred in many developed countries.
